---
title: 'Lecture6 Notes: Tidy data and iteration'
author: "Wenjie Qi"
date: "2021/10/10"
output: html_document
---


```{r}
library(dplyr)
library(palmerpenguins)
library(tidyverse)
library(tidyr)
library(english)
```

# Tidy data


## What is tidy data?

* Each row corresponds to a specific and unique observation representing a similar sort of thing.
* Columns correspond to variables with the same sort of value in each row.

Tidy data is typically far easier to manipulate and apply statistical analysis to in R.


## Reshaping data

```{r}
species <- as.factor(c("Adelie", "Adelie", "Adelie", "Chinstrap", "Chinstrap", "Chinstrap", "Gentoo", "Gentoo", "Gentoo"))
property <- c("bill", "flipper", "weight", "bill", "flipper", "weight", "bill", "flipper", "weight")
value <- c(38.8, 190., 3701., 48.8, 196., 3733., 47.5, 217., 5076.)

penguins_summary_narrow = as_tibble(data.frame(species, property, value))
penguins_summary_narrow
```


### pivot_wider()
```{r}
penguins_summary_narrow <- penguins_summary_narrow %>%
  pivot_wider(names_from = property, values_from = value)  # 3 arguments: df, names_from, value_from

penguins_summary_narrow
```


### pivot_longer()
```{r}
penguins_summary_narrow %>%
  pivot_longer(cols = c("bill", "flipper", "weight"), names_to = "property", values_to = "value")
```

#### a more efficient way
```{r}
penguins_summary_narrow %>%
  pivot_longer(cols = !species, names_to = "property", values_to = "value")
```



## Pivoting for tidy data

```{r}
name <- c("Alice", "Bob", "Charlie")
`2018` <- c(5., 9., 3.)
`2019` <- c(9., 7., 2.)
wins_df_wide <- as_tibble(data.frame(name, `2018`, `2019`))

`2018` <- c(7., 15., 12.)
`2019` <- c(8., 4., 10.)
losses_df_wide <- as_tibble(data.frame(name, `2018`, `2019`))

wins_df_wide
losses_df_wide
```

### paivot_longer()
```{r}
wins_df_narrow <- wins_df_wide%>%
  pivot_longer(!name, names_to = "year", values_to = "wins")
wins_df_narrow

losses_df_narrow <- losses_df_wide %>%
  pivot_longer(!name, names_to = "year", values_to = "losses")
losses_df_narrow
```

```{r}
wins_losses_df <- inner_join(wins_df_narrow, losses_df_narrow) %>%
  mutate(win_rat = wins/ (wins + losses))

wins_losses_df
```


## Uniting and separating data

```{r}
name <- c("Alice", "Alice", "Bob", "Bob", "Charlie", "Charlie")
year <- c("2018", "2019", "2018", "2019", "2018", "2019")
w_over_t <- c("5/12", "9/17", "9/24", "7/11", "3/15", "2/12")
wins_over_total_df <- as_tibble(data.frame(name, year, w_over_t))
wins_over_total_df
```


### separate()
```{r}
sep_df <- wins_over_total_df %>%
  separate(w_over_t, into = c("wins", "totals"), sep = "/", convert = T)  # convert = T
sep_df
```


```{r}
sep_df %>%
  mutate(losses = totals - wins) %>%
  select(-totals)
```


### unite()
```{r}
uni_df <- sep_df %>%
  unite(w_over_t, wins, totals, sep="/")
uni_df
```




# Iteration

## map()
```{r}
is_div_2_3 <- function(x){
  if(x %% 2 == 0){
    return(T)
  }else if(x %% 3 ==0){
    return(T)
  }else{
    return(F)
  }
}
is_div_2_3(3)
```

```{r}
v <- c(1, 2, 3, 4, 5)
is_div_2_3(v)
```

```{r}
map(v, is_div_2_3)
```

### map_lgl()
```{r}
map_lgl(v, is_div_2_3)
```

### map_dbl()
```{r}
example_f <- function(x){
  if(is_div_2_3(x)){
    return(x)
  }else{
    return(0)
  }
}

map_dbl(v, example_f)
```

### map_chr()
```{r}
example_f_eng <- function(x){
  as.character(as.english(example_f(x)))
}

map_chr(v, example_f_eng)
```



## Vectorization
```{r}
is_div_2_3_vect <- function(x){
  return(x %% 2 == 0 | x %% 3 == 0)
}
is_div_2_3_vect(v)
```

```{r}
example_f_vect <- function(x){
  return(x * is_div_2_3_vect(x))
}
example_f_vect(v)
```

```{r}
#example_f_eng_vect <- function(x){
#  return(as.english(example_f_eng_vect(x)))
#}
#example_f_eng_vect(v)
```


# Nesting and unnesting

```{r}
name <- c("Mick", "John", "Paul", "Keith")
band <- c("Stones", "Beatles", "Beatles", NA)
plays <- c(NA, "guitar", "bass", "guitar")
musicians <- data.frame(name, band, plays)
musicians
```

## nest()

```{r}
musicians_nested <- musicians %>%
  group_by(name) %>%
  nest()
musicians_nested
```

## unnest()
```{r}
musicians_unnested <- musicians_nested %>%
  unnest(cols = data)
musicians_unnested
```


# map() & nest()

## create a function which:
*   takes as input a data frame and variable name
*   computes the correlation with all other numeric variables
*   return the name of the variable with maximal absolute correlation, and the correspongding correlation

```{r}
max_cor_var <-function(df, col_name){
  
  v_col <- df %>%
    select(all_of(col_name))  # extract variable based on col_name
  
  df_num <- df %>%
    select_if(is.numeric) %>%
    select(-all_of(col_name))  # select all numeric variables excluding col_name
  
  correlations <- unlist(map(df_num, function(x){
    cor(x, v_col, use = "complete.obs")
  }))  # compute correlations with all other numeric variables

  
  max_abs_cor_var <- names(which(abs(correlations) == max(abs(correlations))))  # extract the variable name
  cor <- as.double(correlations[max_abs_cor_var])  # compute the correlation
  
  return(data.frame(var_name = max_abs_cor_var, cor = cor))  # return dataframe
}

penguins %>%
  max_cor_var("body_mass_g")
```


## combine nest() and unnest() to compute the variable with maximal correlation
```{r}
penguins %>%
  group_by(species) %>%
  nest() %>%
  mutate(max_cor = map(data, ~max_cor_var(.x, "body_mass_g"))) %>%   # ?????????????????????????????????? ??????
  select(-data) %>%
  unnest(cols = max_cor)
```



# Missing data and map()

1. Explicit missing data: The value of an individual variable replaced with "NA".
2. Implicit missing data: Entire rows missing from a data frame.

```{r}
name <- c("Alice", "Charlie", "Alice", "Bob", "Charlie")
year <- c(2018, 2018, 2019, 2019, 2019)
wins <- c(5, 9, 3, 9, 7)
losses <- c(7, NA, 12, 8, 4)

w_l <- data.frame(name, year, wins, losses)
w_l
```


## complete(): make implicit missing data into explicit missing data
```{r}
w_l_na <- w_l %>% 
  complete(name, year)
w_l_na
```


## complete.cases()
get the rows for which there are no NAs, but can lead to reduction in the number of rows and a substantial loss of information
```{r}
complete.cases(w_l)
```

```{r}
w_l %>%
  filter(complete.cases(.))  
```


## Imputation by mean($\mu$)

```{r}
impute_by_mean <- function(x){
  
  mu <- mean(x, na.rm=1) # first compute the mean of x
  
  impute_f <- function(z){  # coordinate-wise imputation
    if(is.na(z)){
      return(mu)  # if z is na replace with mean
    }else{
      return(z)  # otherwise leave in place
    }
  }
  
  return(map_dbl(x, impute_f))  # apply the map function to impute across vector
}
```

```{r}
x <- c(1, 2, NA, 4)
impute_by_mean(x)
```

```{r}
w_l_na
```


```{r}
w_l_na %>%
  mutate(wins = impute_by_mean(wins), losses = impute_by_mean(losses))
```