---
title: "Lecture4 Notes: Data Wrangling"
author: "Wenjie Qi"
date: "2021/10/3"
output: html_document
---

## Tabular data

```{r}
library(tidyverse)
library(palmerpenguins)
```

Rows: examples, observations, cases
Columns: features, variables, covariates
```{r}
head(penguins)
```

# Data Wrangling
Data wrangling is the process of transforming data from one form to another
Extracting, transforming, fusing and aggregating information from exist data
We can do this all in R with the Tidyverse, especialling the dplyr

### "nouns"
  Data frames, the representations in R of tabular data
  
### "verbs"
  Operations which can be applied to data frames
  select()   - take a subset of columns 
  filter()   - take s subset of rows
  mutate()   - add or modify existing columns
  arrange()  - sort rows
  summarize()- aggregate data across existing rows

## select()
extract a subset of columns
```{r}
select(penguins, species, bill_length_mm, body_mass_g)
```

remove several columns: put a minus before arguments
```{r}
select(penguins, -species, -bill_length_mm, -body_mass_g)
```

## filter()
extract a subset of rows
```{r}
filter(penguins, species=="Gentoo")
```

we can combine two or more conditions
```{r}
filter(penguins, species=="Gentoo" & body_mass_g>5000)
```
  
we can combing select() and filter()
```{r}
select(filter(penguins, species=="Gentoo"), species, bill_length_mm, body_mass_g)
```


## The pip operator "%>%"
chaining multiple operations with the pip operator %>%
```{r}
penguins %>%
  filter(species=="Gentoo") %>%
  select(species, bill_length_mm, body_mass_g)
```

%>% allow arguments to be implicitly passed as object to the function after the pip
```{r}
f <- function(a,b){
  return(a^2 + b)
}
f(3,1)
```

```{r}
3 %>% f(1)  # "3" is supplied as the first argument in the function
```

## mutate()
create a new column as a funtion of existing columns
```{r}
my_penguins <- penguins %>%
  mutate(flipper_bill_ratio = flipper_length_mm/bill_length_mm) %>%
  select(species, bill_length_mm, flipper_length_mm, flipper_bill_ratio)
my_penguins
```

```{r}
ggplot(data=rename(my_penguins, Species=species), aes(x=flipper_bill_ratio, y=Species, fill=Species)) + geom_violin() + theme_bw() + xlab("Flipper bill ratio")
```


## rename()
```{r}
my_penguins %>% rename(f_over_b = flipper_bill_ratio)
```


## arrange()
sort the rows of a table via arrange()
```{r}
my_penguins %>% arrange(desc(bill_length_mm))
```

  
## summarize()
summarize() computes vector functions across the entire data frame
```{r}
penguins %>%
  summarize(
    num_rows=n(), avg_weights_kg=mean(body_mass_g/1000, na.rm=T), avg_flipper_bill_ratio=mean(flipper_length_mm/bill_length_mm, na.rm=T)
  )
```

combine summarize() and group_by()
```{r}
penguins %>%
  group_by(species) %>%
  summarize(
    num_rows=n(), avg_weight_kg=mean(body_mass_g/1000,na.rm=T), avg_flipper_bill_ratio=mean(flipper_length_mm/bill_length_mm,na.rm=T)
  )
```


## across()
apply a function within summarize to all columns at once
```{r}
# summarize(penguins, across(everything(), ~sum(is.na(.x))))
penguins %>% 
  summarize(across(everything(), ~sum(is.na(.x))))
```


## across() with where()
```{r}
penguins %>%
  summarize(across(where(is.numeric), ~mean(.x, na.rm=T)))
```


## Combining summarize(), group_by() and across()
```{r}
penguins %>%
  select(-year) %>%
  group_by(species) %>%
  summarize(across(where(is.numeric), ~mean(.x, na.rm=T)), num_rows=n())
```


## join()
fuse multiple data frames
```{r}
penguin_bill_lengths_df <- penguins %>%
  arrange(desc(bill_length_mm)) %>%
  select(species, bill_length_mm)
penguin_bill_lengths_df
```

```{r}
species <- unique(penguins$species)
latin_name <- c("Pygoscelis adeliae", "Pygoscelis papua", "Pygoscelis antarcticus")
latin_name_df <- data.frame(species, latin_name)
latin_name_df
```

```{r}
penguin_bill_lengths_df %>%
  inner_join(latin_name_df)
```


###  Types of join function
1.    inner join: only fuse the rows appears in both data frames
2.    full join: opposite of inner join
3.    left join: contain all rows of the left data frame
4.    right join: opposite of left join




