---
title: "Lecture3 Notes: Visualisition with R"
author: "Wenjie Qi"
date: "2021/10/2"
output: html_document
---



# The importance of visualisition
1.    Exploring data
2.    Communicating your insights



# Visual cues
1.    Position
2.    Length
3.    Angle
4.    Direction
5.    Shape
6.    Area
7.    Volume
8.    Shade
9.    Color



# Visualisition in R with ggplot2

##use install.packages("tidyverse") to install ggplot2, dplyr, tidyr, purrr

```{r}
library(ggplot2)
library(palmerpenguins)
library(dplyr)  # to use rename
```


## Types of variables

Continuous:  Numeric variables that can take any value on an interval
  e.g. Bill length, bill depth, flipper length, body mass
  
Discrete:  Numeric variables for which there is a minimum gap between possible values
  e.g. year the observation was recorded

Categorical:  Variables that can take on only a specific set of values representing distinct categories
  e.g. species, island, etc.
```{r}
head(penguins)  # show the head rows of penguins data set
```


## Univariate plots

### Histogram plot
```{r}
univar_plot <- ggplot(data=penguins, aes(x=flipper_length_mm))+xlab("Flipper length (mm)")
univar_plot + geom_histogram(binwidth=5) + ylab("Count")
```


### Kernel density plot
```{r}
univar_plot + geom_density(adjust=0.5) + ylab("Density")
```


###  Skewness
Negative skewed data: a large left tail consisting of a relatively small number of relatively low values, but most of the data is towards the upper end of the plot
Positive skewed data: a large right tail
Can be seen from a density plot


### Unimodal vs. multi-modal
The number of modes refers to the number of peaks within the data.
So the univar_plot above is a bimodal distribution.


## Bivariate plots

### Density plot
```{r}
ggplot(data=rename(penguins, Species=species), aes(x=flipper_length_mm, color=Species)) + geom_density() + theme_bw() + xlab("Flipper length (mm)") + ylab("Density")
```


### Box plot
```{r}
ggplot(data=penguins, aes(x=flipper_length_mm, y=species)) + geom_boxplot() + xlab("Flipper length (mm)") + ylab("Penguin species")
```


### Violin plot
```{r}
ggplot(data=rename(penguins, Species=species), aes(x=flipper_length_mm, y=Species, fill=Species)) + geom_violin() + theme_bw() + xlab("Flipper length (mm)")
```


### Scatter plot: to show the relationship between two continous variables
```{r}
mass_flipper_scatter <- ggplot(data=penguins, aes(y=body_mass_g, x=flipper_length_mm)) + xlab("Flipper length (mm)") + ylab("Body mass (g)")
mass_flipper_scatter + geom_point(size=2)
```



## Multivariate plots

### Scatter plot
```{r}
mass_flipper_scatter + geom_point(aes(color=bill_length_mm), size=2) + scale_color_gradient(low="blue", high="red") + guides(color=guide_legend("Bill length (mm)"))
```


```{r}
mass_flipper_scatter + geom_point(aes(color=bill_length_mm, size=bill_depth_mm)) + scale_color_gradient(low="blue", high="red") + guides(color=guide_legend("Bill length (mm)"), size=guide_legend("Bill depth (mm)"))
```


when use color, use an additional aesthetic to be clear
```{r}
mass_flipper_scatter + geom_point(aes(color=species, shape= species))
```


```{r}
mass_flipper_scatter + geom_text(aes(label=species, color=species)) + guides(color=guide_legend("Species"))
```

### Facets
```{r}
mass_flipper_scatter + geom_point() + facet_wrap(~species)
```



## Trend lines: illustrate the relationship between two variables
```{r}
trend_plot <- ggplot(data=filter(penguins, species=="Gentoo"), aes(y=body_mass_g, x=flipper_length_mm)) + xlab("Flipper length (mm)") + ylab("Body mass (g)") + geom_point()
trend_plot + geom_smooth()
```


```{r}
trend_plot + geom_smooth(method="lm")
```


```{r}
min(filter(penguins, species=="Gentoo")$body_mass_g,na.rm=T)
```



## Annotation
```{r}
trend_plot + geom_smooth(method="lm") + geom_curve(x=220, xend=209, y=4250, yend=3975, arrow=arrow(length=unit(0.5, "cm")), curvature=0.2) + geom_text(x=225, y=4250, label="The lightest Gentoo \n penguin wighs 39.5kg")
```

