---
title: "SCEM_Assignment_03"
author: "Wenjie Qi"
date: "2021/10/13"
output: html_document
---


# 1.  Random experiments, events and sample spaces

* Random experiment:
a procedure which:
  (a) has a well-defined set of possible outcomes;
  (b) could (at least in principle) be repeated arbitrarily many times.


* Event:
a set of possible outcomes.

* Sample space:
the set of all possible outcomes of interest for a random experiment.

# 2. Tidy data and iteration
## 2.1 Missing data and iteration

```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(palmerpenguins)

data("penguins")
```


```{r}
impute_by_mean <- function(x){
  mu <- mean(x, na.rm = 1)  # first compute the mean of x
  
  impute_f <- function(z){  # coordinate-wise imputation
    if(is.na(z)){
      return(mu)
    }else{
      return(z)
    }
  }
  
  return(map_dbl(x, impute_f))  #apply the map function to the impute across vector
}
```

```{r}
impute_by_median <- function(x){
  med <- median(x, na.rm=1)
  
  impute_f <- function(z){
    if(is.na(z)){
      return(med)
    }else{
      return(z)
    }
  }
  
  return(map_dbl(x, impute_f))
}

v <- c(1, 2, NA, 4)
impute_by_median(v)
```

```{r}
x <- c(seq(0, 10, 0.1))
x
y <- 5 * x + 1

df_xy <- data.frame(x, y)
df_xy %>% head()
```

```{r}
df_xy %>% 
  mutate(z = map2_dbl(x, y, ~.x+.y)) %>%
  head(5)
```


## df with missing data

```{r}
sometimes_missing <- function(index, value){
  if(index %% 5 == 0)
    return(NA)
  else
    return(value)
}
```


```{r}
x <- seq(0, 10, 0.1)
df_xy_missing <- as.tibble(data.frame(x))

df_xy_missing <- df_xy_missing %>%
  # mutate(y_missing = map2_dbl(x, y, ~sometimes_missing(row_number(), .y)))
  mutate(y = map2_dbl(.x=row_number(), .y=y, ~sometimes_missing(.x, .y)))
df_xy_missing
```

```{r}
# x <- seq(0, 10, 0.1)
# y_missing <- c()
# for(i in 1:101){
#   if(i%%5==0)
#     y_missing[i] <- NA
#   else
#     y_missing[i] <- 5*x[i]+1
# }
# df_xy_missing <- data.frame(x, y_missing)
# df_xy_missing %>% head(10)
```

```{r}
x <- seq(0, 10, 0.1)
df_xy_imputed <- data.frame(x)
df_xy_imputed <- df_xy_imputed %>%
  mutate(y = impute_by_median(df_xy_missing$y))
df_xy_imputed
```


```{r}
#names(df_xy_missing)[names(df_xy_missing) == "y_missing"] <- "y"
#names(df_xy_imputed)[names(df_xy_imputed) == "y_imputed"] <- "y"

df_xy <- df_xy %>%
  mutate(source="original") 

df_xy_missing <- df_xy_missing %>%
  mutate(source="corrupted")

df_xy_imputed <- df_xy_imputed %>%
  mutate(source="imputated")

df_combined <- combine(df_xy, df_xy_missing, df_xy_imputed)
```

```{r}
ggplot(df_combined, aes(x=x, y=y, color=source)) +geom_point() + facet_wrap(~source) + geom_smooth(method="lm")
```


## 2.2 Tidying data with pivot functions

```{r}
library(readxl)
```

```{r}
folder_path <- "D:\\Develop\\Projects\\R\\SCEM\\Lectures\\Assignments\\"

file_name <- "HockeyLeague.xlsx"

file_path <- paste(folder_path, file_name, sep="")

wins_data_frame <- read_excel(file_path, sheet="Wins")
losses_data_frame <-read_excel(file_path, sheet="Losses")
```


```{r}
wins_data_frame %>%
  select(1:5) %>% 
  head(3)
```

```{r}
wins_tidy <- wins_data_frame %>%
  rename(Team = ...1) %>%
  pivot_longer(!Team, names_to = "Year", names_transform = list(Year = as.integer), values_to = "w_over_t") %>%
  separate(w_over_t, into = c("Wins", "Total"), sep = " of ", convert = T)

wins_tidy %>% dim()
wins_tidy %>% head(5)
```



```{r}
losses_data_frame %>%
  select(1:5) %>%
  head(3)
```

```{r}
losses_tidy <- losses_data_frame %>%
  rename(Team = ...1) %>%
  pivot_longer(!Team, names_to = "Year", names_transform = list(Year = as.integer), values_to = "l_over_t") %>%
  separate(l_over_t, into = c("Losses", "Total"), sep = " of ", convert = T)

losses_tidy %>% dim()
losses_tidy %>% head(5)
```


```{r}
hockey_df <- wins_tidy %>%
  inner_join(losses_tidy) %>%
  mutate(Draws = Total - Wins - Losses) %>%
  mutate(across(!c(Team, Year, Total), ~ .x / Total, .names = "{col}_rt"))

hockey_df
```


### Make a summary
```{r}
hockey_df$Team <- as.factor(hockey_df$Team)

hockey_summary <- hockey_df %>%
  select(Team, Wins_rt, Losses_rt, Draws_rt) %>%
  group_by(Team) %>%
  summarise(Mean_Wins_rt=mean(Wins_rt),Median_Wins_rt=median(Wins_rt), Mean_Losses_rt=mean(Losses_rt), Median_Losses_rt=median(Losses_rt), Mean_Draws_rt=mean(Draws_rt), Median_Draws_rt=median(Draws_rt)) %>%
  arrange(desc(Mean_Wins_rt))

hockey_summary
```


## 2.3 Most correlated variables (*)

```{r}
max_cor_var <- function(df, col_name){
  
  v_col <- df %>%
    select(all_of(col_name))  # extract variable based on col_name
  
  df_num <- df %>%
    select_if(is.numeric) %>%
    select(-all_of(col_name))  # select all numeric variables excluding col_name
  
  correlations <- unlist(map(df_num, function(x){
    cor(x, v_col, use = "complete.obs")
  }))  # compute correlations with all other numeric variables

  
  max_abs_cor_var <- names(which(abs(correlations) == max(abs(correlations))))  # extract the variable name
  
  
  return(max_abs_cor_var)  # return dataframe
}


penguins %>%
  max_cor_var("body_mass_g")
```


```{r}
top_correlates_by_var <- function(df){
  
  cols <- df %>%  # list of all numeric variables
    select_if(is.numeric)  
  
  col_names <- names(cols) # list of all those variables' names
  
  max_cor_var_names <- list()  # make a empty list to contain max_cor_var_names
  
  index = 1
  for(var in col_names){
    max_cor_var_names[index] <- max_cor_var(df, var)  # map()
    index <- index + 1
  }
  
  return_df <- as.tibble(data.frame(max_cor_var_names))  # turn this list to a dataframe, but the name of columns are not what we want
  
  names(return_df) <- col_names  # rename the columns' names with col_names
  
  return(return_df)
}

penguins %>%
  top_correlates_by_var()

```


```{r}

penguins %>%
  group_by(species) %>%
  nest() %>%
  mutate(top_cor = map(data, top_correlates_by_var)) %>%
  select(-data) %>%
  unnest(cols = c(top_cor)) %>%
  select(-year)
  
```



# 3.5 (**)

Proof:

Assume that [1,0] is countable finite.

Then we can display all the numbers belong to (0, 1) in this format:

0.a11 a12 a13...
0.a21 a22 a23...
0.a31 a32 a33...
....
0.an1 an2 an3...
...

then we can construct a number:
0.ax1 ax2 ax3...
and let ax1!= a11, ax2 != a22, ax3 != a33,... , axn != ann,...