---
title: "SCEM_Assignment_2"
author: "Wenjie Qi"
date: "2021/10/6"
output: html_document
---

```{r}
library(tidyverse)
library(Stat2Data)
library(ggplot2)

data("Hawks")
hawksSmall <- drop_na(select(Hawks, Age, Day, Month, Year, CaptureTime, Species, Wing, Weight, Tail))
```

# 1 Visualisation

## 1.1 Types of variables
```{r}
dim(hawksSmall)
```

```{r}
head(hawksSmall)
```

1. Month:    discrete
2. Species:  categorical
3. Age:      discrete
4. Wing:     continuous
5. Weight:   continuous


## 1.2 What's wrong with this plot?
1. There's no labels of horizontal and vertical direction.
2. The points are to large to see clearly.
3. The guide legend which says 'Culmen' cannot make us know which attribute or variable of culmen it represents.
4. There are too many variables in that scatter plot, it's hard to read.


## 1.3 Generate a histogram
```{r}
weight_plot <- ggplot(data = hawksSmall, aes(x = Weight)) + xlab("Weight (gm)")
weight_plot + geom_histogram(binwidth = 100) + ylab("Count")
```

